6.867 Problem Set 1

### 2015-12-10 Project/Latex/Paper.tex
[project] dave revise second half.

Once we have constructed our full set of features, we aim to predict log(article pageviews), $\mathbf{y}$, using our full set of variables, which we include in our design matrix, $\mathbf{X}$. We estimate the feature weights using the closed form solution for both OLS and ridge regression:
\begin{equation}
\beta = (\mathbf{X}^T\mathbf{X} + \lambda \mathbf{I})^{-1}\mathbf{X}^T\mathbf{y}
\end{equation}
where $\mathbf{I}$ is the $k \times k$ identity matrix, and $\lambda$ is our regularization parameter. Setting $\lambda = 0$ corresponds to OLS, whereas non-zero values of $\lambda$ correspond to ridge regression. The motivation for performing ridge regression as opposed to OLS is to not overfit on our data \cite{ridgeregressionNotes}.

In order to choose an appropriate value of $\lambda$, we split our data into a training and validation set. 90\% of the data is allocated to the training set, and 10\% of the data is allocated to the test set. We cross-validate the estimated $\beta$ values on our validation set for each $\lambda$, and choose the value of $\lambda$ that produces the lowest MSE on the validation data.

The table below displays the 20 feature weights with the largest magnitudes given our model. A bar chart showing the magnitude of all of the weights (excluding the intercept weight) can be found in Figures \ref{fig:lin_reg_weights1} and \ref{fig:lin_reg_weights2}.

@@

\begin{figure*}[ht]\centering
\includegraphics[width=\linewidth]{feature_weights.png}
\caption{Linear Regression Feature Weights (excluding intercept term)}
\label{fig:lin_reg_weights1}
\end{figure*}

\begin{figure*}[ht]\centering
\includegraphics[width=\linewidth]{feature_weights_contd.png}
\caption{Linear Regression Feature Weights (excluding intercept term)}
\label{fig:lin_reg_weights2}
\end{figure*}

Figure \ref{fig:mse_plot} shows the training and holdout MSE for various values of Lambda. There are a few things in this plot worth discussing. First, note that the holdout MSE is consistently lower than the training data MSE. Given the (relatively) small size of our dataset (6,687 observations), this is likely due to the sampling we used to separate our data into training and validation data. However, this shouldn't effect the validity of our cross validation.

Another thing worth noting is that even on the training dataset, the MSE goes down once we choose a non-zero value of $\lambda$. At first, this was surprising to our group, as conceptually OLS is often thought of as the linear regression method that minimizes MSE. However, it is important to note that OLS only holds this distinction amongst unbiased estimators. Hoerl and Kennard (1970) \cite{hoerl1970ridge} prove the existence theorem for ridge regression, which proves that there exists a value of $\lambda$ for which $\beta_{ridge}$ produces a lower MSE than $\beta_{OLS}$.

@@

\noindent For some values of $\lambda$, ridge regression is able to lower the MSE by decreasing variance, but adding a non-zero bias term. We believe this is what is happening in our data when cycling through different $\lambda$ values.

@@

@@

Given the large amount of work we put into building numerous text-based features (such as the trigram neural network perplexity, the sentiment, and the Flesch reading ease), we may want to evaluate how impactful these various features actually are in our regression. How much incremental reduction in MSE are we getting by including them? We first re-run the exact same regression specification, but instead of using the trigram perplexity calculated from our neural network, we instead use the simple bigram perplexity discussed earlier in this paper. We find that this actually causes the MSE to decrease, from 2.392 to 2.389. This small, but modest change suggests that there is almost no incremental value from the neural network feature.

Now, we might ask if any text features are adding much value. Particularly given that, in general, the magnitude of their weights is dwarfed by the magnitude of the weights of the contextual features when looking at feature weights, we might expect that text features do not add much. When removing perplexity, sentiment, reading ease, and word count features from our regression, we find that the MSE increases, from 2.392 to 2.502. This is result is encouraging, as it suggests that even if text features aren't doing much, they're doing something! Removing them from our model leads to a 4.6\% increase in the MSE.

@@

\section{Future Work}
Unfortunately due to time constraints as well as various unanticipated difficulties and technical errors, we were unable to implement all of several important features we had initially wanted to include in our regression. Most notably, this paper in its current state lacks some form of topic modeling which we expect to be a strong predictor of content viewership at least in our context. We also think the constructing some features using only the headlines might also provide some improvement, especially considering that an article's headline usually ends up being fairly important in a user's decision to click through or not.

There is also much room for improvement with our existing features as well. For example, the perplexity score feature ended providing very little predictive power might be a result of the underlying language model not being good enough, especially considering that the neural network model still hasn't completed its training. Though we did observe that there was virtually no difference moving from the bigram perplexity scores to the trigram neural network perplexity scores, this again might be because the neural network wasn't given sufficient time to train. We can imagine that if we continued training our neural network for several more cycles, we might start to see visible performance gains. Alternatively, we think its likely the its not the perplexity of an article relative to the entire corpus that matters, but rather perplexity relative to other articles that cover the same event or topic.

Similarly, there is ample room to improve our sentiment analysis methodologies. The training dataset that we generated using mechanical turk is relatively small (200 articles). Because of this, there are currently many words that may have strong conditional probabilities of appearing given a particular article sentiment that simply do not appear in our training data. With more time (and money!), we might be able to expand the size of our training data and consequently include the accuracy of our model. Furthermore, the model currently skips words it hasn't seen in the training corpus - it might make sense to augment our Naive Bayes implementation to incorporate a method such as Laplace smoothing so as to not simply ignore words we haven't seen in our training data.

Furthermore, the current discrepancies between our implementation of Naive Bayes and NLTK's, while not large in magnitude, are alarming. Given more time in the future, we hope to dig deeper into what is happening. Our current hypothesis is that NLTK's implementation of Naive Bayes does not take into account the number of times a particular word appears in a document. Because of this, more distinct informative words are required to move a given article's prediction away from the neutral class, which has an extremely strong prior.

Another way we could improve our model is to apply a basis expansion to our variables to include polynomial and various interaction terms. In many cases, theres no good justification to why a feature is linearly related to the output feature. Hence, applying a polynomial expansion might uncover an entirely different relationship between our dependent variable and its covariates. We also think that interaction terms might provide some significant predictive power.

Certainly one thing we definitely plan on doing if we end up trying to turn this project into a publishable paper is to include all of the articles in our dataset. With the much greater sample size and the much more expansive corpus from the additional articles, we might expect to see drastically different optimal weights.

Since we also have Twitter data that tracks every tweet and retweet involving a NYT article, we would like to extend this project to also include our Twitter data. We can then explore many interesting potential cross effects, such as how does the article content predict the amount of Twitter sharing and alternately, how does content of a Tweet predict viewership.

@@

The work for this project was divided as follows. Michael extracted the New York Times contextual data and page view counts and wrote the NYT website scraping script. Dave and Jeremy built out many of the content-driven features and additional contextual features, such as word count, author gender, popularity, and Flesch reading ease. Dave and Jeremy built the implementation and validation of the Naive Bayes classifier, while Michael focused his efforts on the construction and validation of the language model. Dave, along with assistance from Michael, focused on the implementation of OLS and ridge regression, along with validation and discussion of results. Plot generation, table creation, writing, and editing was evenly split between Michael and Extenuating personal circumstances unfortunately prevented Jeremy from contributing to the project as much as originally anticipated.

>>>>

Using our full set of features, we are now able to perform the regression task we originally had in mind, and attempt to determine how (if at all) content drives viewership. We regress log(article pageviews), $\mathbf{y}$, on our design matrix, $\Phi$, which includes entries for each of our $k-1$ features, plus an intercept term (in this case, $k$ = 102).. We estimate the feature weights using the closed form solution for OLS and ridge regression:
\begin{equation}
\beta = (\Phi^T\Phi + \lambda \mathbf{I})^{-1}\Phi^T\mathbf{y},
\end{equation}
\noindent where $\mathbf{I}$ is the $k \times k$ identity matrix, and $\lambda$ is our regularization parameter. Setting $\lambda = 0$ corresponds to OLS, whereas a non-zero value of $\lambda$ corresponds to ridge regression. The motivation for performing ridge regression as opposed to OLS is to not overfit on our data, and the value of $\lambda$ can be interpreted as the strength of our Bayesian prior on the feature weights being equal to 0 \cite{ridgeregressionNotes}.

In order to choose an appropriate value of $\lambda$, we split our data into training, validation, and test sets. 90\% of the data is allocated to the training set, 10\% to the validation set, and 10\% to the test set. Although we estimate $\beta$ on the training data using the above closed-form solution, we cross-validate on our validation set for each $\lambda$, and choose the value of $\lambda$ that produces the lowest MSE on our validation dataset. To ensure that we have not overfit $\beta$ to our validation dataset, we also calculate the MSE on the test set as a final step. A comparison of the training, validation, and test MSEs for various values of $\lambda$ is found in Figure \ref{fig:mse_plot}.

We find that $\lambda = 100$ minimizes the MSE on our validation set. Table 6 displays the 20 feature weights with the largest magnitudes $\lambda = 100$. Two charts showing weights for the full set of features (excluding the intercept term) can be found in Figures \ref{fig:lin_reg_weights1} and \ref{fig:lin_reg_weights2}.

@@

\begin{figure*}[ht]\centering
\includegraphics[width=\linewidth]{feature_weights.png}
\caption{Linear Regression Feature Weights (excluding intercept term) (1 of 2)}
\label{fig:lin_reg_weights1}
\end{figure*}

\begin{figure*}[ht]\centering
\includegraphics[width=\linewidth]{feature_weights_contd.png}
\caption{Linear Regression Feature Weights (excluding intercept term) (2 of 2)}
\label{fig:lin_reg_weights2}
\end{figure*}

It's worth taking the time to discuss Figure \ref{fig:mse_plot}, which shows the training and holdout MSE for various values of $\lambda$, in slightly more depth. There are a few things in this plot worth discussing. First, note that the validation MSE is consistently higher than the training data MSE, which is consistently higher than the test data MSE. Given the (relatively) small size of our dataset (6,687 observations), this is likely due to the sampling we used to separate our data into training, validation, and test data. However, we don't expect this effect the validity of our cross validation.

Another thing worth noting is that even on the training dataset, there exist non-zero values of $\lambda$ that achieve a lower MSE than the OLS estimate of $\beta$. At first, this was surprising to our group, as conceptually OLS is often thought of as the linear regression method that minimizes MSE. However, it is important to note that OLS only holds this distinction amongst unbiased estimators. Hoerl and Kennard (1970) \cite{hoerl1970ridge} prove the existence theorem for ridge regression, which claims the existence of some $\lambda$ such that $\beta_{ridge}$ produces a lower MSE than $\beta_{OLS}$.

@@

\noindent For some values of $\lambda$, ridge regression is able to lower the MSE by decreasing variance, but increase the bias from zero to some non-zero value. We believe the changes in MSE we observe in our dataset as we vary $\lambda$ can be explained by this phenomenon.

@@

In general, we can now interpret the strength of the feature weights produced by our regression to determine how predictive a particular feature is of readership. Note from Table 6 that most of the features with the most predictive power are not the text-based features. The intercept term in our regression is orders of magnitude larger than any other feature, implying that most of the NYTimes articles receive many pageviews in the baseline case. The strongest coefficients tend to be those that indicate the publication desk, section, and time of publication of the article. This suggests that the content of an article itself may not be as important as the context in which it is published. The one exception we see is that a higher word count is predictive of higher viewership. We suspect that what's going on here is correlative, rather than causal - the quality of longer pieces (e.g., NYTimes Magazine articles) is likely higher, thus driving more readership. But we don't expect that a website full of low-quality, 5,000 word pieces would be successful.

@@

Given the large amount of work we put into building numerous text-based features (such as the trigram neural network perplexity, the sentiment labels, and the Flesch reading ease) and the relatively low impact they seem to have had on our regression (based on feature weights), we want to specifically evaluate the impact of these features on our regression. Specifically, how much incremental reduction in MSE are we getting by including them? We first re-run the exact same regression specification, but instead of using the trigram perplexity calculated from a neural network, we instead use the simple bigram perplexity discussed earlier in this paper. We find that this actually leads to a reduction in MSE, from 2.392 to 2.389. This small, but modest change suggests that there is currently almost no incremental value from using a neural network language model as opposed to a more straightforward language model.

As a next step, we ask if text features in general add much value to our model. Given that, in general, the magnitude of text feature weights is dwarfed by the magnitude of contextual feature weights, we might expect that text features do not add much. We find that removing the perplexity, sentiment, reading ease, and word count features from our regression leads to an increase in MSE, from 2.392 to 2.502. This result is encouraging, as it suggests that even if text features currently aren't contributing as much as we had hoped, they are doing something! Removing them from our model leads to a 4.6\% increase in the MSE.

@@

\section{Future Work}
Unfortunately, we were unable to implement all of several important features we initially wanted to include in our regression, due to a combination of time constraints and various unanticipated technical difficulties and errors. Most notably, this paper currently lacks some form of topic modeling, which we expect to be a strong predictor of content viewership (at least in our context). We also suspect that the addition of features containing information about article headlines may provide some improvement. It seems reasonable to assume that an article's headline is a fairly important factor in a user's decision to click through or not.

There is also room for improvement in the design and extraction of our existing features. For example, the perplexity score feature ultimately provided very little predictive power. This may be due to the issues in the quality of the underlying language model. This is particularly likely given that the neural network model was not able to complete many iterations, and has not yet converged. Although we observed virtually no difference in performance moving from the bigram perplexity scores to the trigram neural network perplexity scores, this again may be due to insufficient training time for the neural network. We find it conceivable that, given many more cycles to train, our neural network would start to yield visible performance gains over the bigram perplexity score. Another hypothesis is that rather than measuring the the perplexity of an article relative to the entire corpus of articles, it would be more effective to measure a given article's perplexity relative to other articles that cover the same event or topic.

Tshere is ample room to improve our sentiment analysis methodology. The training dataset generated using mechanical turk is relatively small (200 training articles). Because of this, there may exist many words that have a strong probability of appearing in an article conditional on sentiment that simply do not appear in our training set. With more time (and money!), we could label more data and improve the accuracy of our model. Furthermore, the model currently skips words that do not occur in the training corpus. An extension of our Naive Bayes implementation could use a method such as Laplace smoothing, so as to not simply ignore words we haven't seen in our training data.

In addition, the current discrepancies between our implementation of Naive Bayes and NLTK implementation of the algorithm, while not large in magnitude, are alarming. In the future, we hope to dig deeper into this discrepancy and identify the root cause. Our current hypothesis is that NLTK's implementation of Naive Bayes does not take into account the number of times a particular word appears in a document. Because of this, a larger number of distinct informative words are required in order to overcome the strong prior belief that a given article's sentiment is neutral.

Another avenue for potential improvement to our model is the application of basis expansion to our variables. This would allow us to include polynomial and interaction terms. In many cases, there is no good justification for assuming a feature is linearly related to the output feature. Hence, applying a polynomial expansion might uncover an entirely different relationship between our dependent variable and its covariates. By a similar line of reasoning, we suspect that interaction terms may have significant predictive power.

Given that we have access to six full months of NYTimes data, we also hope to expand our dataset to include the full body of NYTimes articles we have available to us. With a larger sample size and more expansive corpus, we might expect to see drastically different optimal weights, as our results will be less sensitive to time-dependent trends in viewership and content. Since we also have access to Twitter data tracking every tweet and retweet involving a NYT article, we also hope incorporate this data into a future version of our model. This would enable us to explore potential cross effects, such as the relationship between article content and Twitter sharing.

@@

The work for this project was divided as follows. Michael extracted the New York Times contextual data and page view counts and wrote the NYT website scraping script. Dave and Jeremy built out many of the content-driven features and additional contextual features, such as word count, author gender, popularity, and Flesch reading ease. Dave and Jeremy built the implementation and validation of the Naive Bayes classifier, while Michael focused his efforts on the construction and validation of the language model. Dave, with assistance from Michael, focused on the implementation of OLS and ridge regression, along with validation and discussion. Plot generation, table creation, writing, and editing was evenly split between Michael and Dave. Extenuating personal circumstances unfortunately prevented Jeremy from contributing to the project as much as originally anticipated - he will likely be in touch soon with the instructors regarding this issue.
